\section{Introduction}
State machine replication protocols are a critical component of many fault
tolerant distributed systems~\cite{burrows2006chubby, corbett2013spanner,
thomson2012calvin, baker2011megastore, taft2020cockroachdb}. Given an arbitrary
deterministic state machine, like a key-value store or a relational database, a
state machine replication protocol can be used to deploy multiple copies, or
replicas, of the state machine while guaranteeing that the states of the
replicas stay in sync and do not diverge.

The most popular and widely deployed state machine replication protocols are
Paxos~\cite{lamport1998part, burrows2006chubby, corbett2013spanner} and
Raft~\cite{ongaro2014search, taft2020cockroachdb, tidb2019website,
yugabyte2019website}. These protocols have two distinguishing characteristics.
First, they are leader based. All communication is funneled through a single
leader. Second, these protocols totally order state machine commands into a log
and have state machine replicas execute the commands in log order. Every
replica executes the exact same commands in the exact same order.

There is another family of \emph{generalized multi-leader} state machine
replication protocols---including EPaxos~\cite{moraru2013there},
Caesar~\cite{arun2017speeding}, and Atlas~\cite{enes2020state}---that improve
the performance of protocols like MultiPaxos and Raft along these two
dimensions. These protocols are \defword{multi-leader} and avoid being
throughput bottlenecked by a single leader. They are also
\defword{generalized}~\cite{lamport2005generalized, losa2016brief}.
% Addresses Reviewer 1.
%
% > In the introduction, I would clarify that this paper's scope is on
% > dependency graph-based Paxos protocols. It
\markrevisions{This means that the protocols are based on dependency graphs.}
Every replica executes non-commuting commands in the exact same order, but the
replicas are free to execute commuting commands in any order. As a result,
commuting commands do not interfere with one another.

Unfortunately, these generalized multi-leader protocols are extremely
complicated. Paxos has a well known reputation for being
complex~\cite{lamport2001paxos, van2015paxos, ongaro2014search}, and these
generalized multi-leader protocols are significantly more complex than that.
They require a strong understanding of more sophisticated Paxos variants like
Fast Paxos~\cite{lamport2006fast} and are overall less intuitive and more
nuanced. It's hard to measure this complexity precisely, but there are
indications that the protocols are complicated. EPaxos, for example, had several
bugs go undiscovered for years despite the popularity of the
protocol~\cite{sutra2011fast}. Through personal conversations, we have also
found that even domain experts find these protocols challenging to fully
understand.

This complexity has negative consequences in industry and academia. The
performance advantages of generalized multi-leader protocols make them an
attractive option for industry practitioners. Despite this, generalized
multi-leader protocols have little to no industry adoption.  We postulate that
this is largely due to their complexity. The engineers
in~\cite{chandra2007paxos} explain that implementing a state machine
replication protocol requires making many small changes to the protocol to
match the environment in which it is deployed. Making these changes without a
strong understanding of the protocol is infeasible. Academically, it is
challenging to compare and contrast the various protocols. They all seem very
similar, yet vaguely distinct. This also makes it difficult to extend the
protocols with further innovations. There are dozens of state machine
replication protocols in the literature, yet relatively few generalized
multi-leader variants.

% Addresses Reviewer 1.
%
% > Finally, I felt the paper wanders among various PPaxos \& other prior work. I
% > wonder if setting up a goal upfront would be helpful. It might be useful to
% > set the goal of understanding EPaxos like protocols and why people want them.
This paper is a tutorial on generalized multi-leader state machine replication
protocols. \markrevisions{Our goal is to} answer questions such as: What
problem do these protocols address? How can I choose between the various
protocols? Why do these protocols work the way they do? What do they have in
common? Where do they differ? Which parts of the protocols are straightforward?
Which are more subtle than they appear? Are there simpler variants out there?
What trade-offs do the protocols make, and which points in the design space are
still unexplored?

The tutorial has four parts, and in each part, we introduce a new protocol.
First, we present the simplest possible generalized multi-leader protocol,
which we called \textbf{Simple \BPaxos{}} (\secref{SimpleBPaxos}). Simple
\BPaxos{} sacrifices performance for simplicity and is designed with the sole
goal of being easy to understand.  Simple \BPaxos{} is the kernel from which
all other generalized multi-leader protocols can be constructed. It
encapsulates all the mechanisms and invariants that are common to the other
protocols.

Second, we introduce a purely pedagogical protocol called \textbf{Fast
\BPaxos{}} (\secref{FastBPaxos}). Fast \BPaxos{} achieves higher performance
than Simple \BPaxos{}, but it is unsafe. The protocol does not properly
implement state machine replication. Why study a broken protocol? Because
understanding why Fast \BPaxos{} does \emph{not} work leads to a fundamental
insight on why other protocols do. Specifically, we discover that generalized
multi-leader protocols encounter a fundamental tension between agreeing on
commands and ordering commands. The way in which a protocol handles this
tension is its key distinguishing feature. We taxonomize the protocols into
those that avoid the tension and those that resolve the tension.

Third, we introduce \textbf{Unanimous \BPaxos{}}, a simple \emph{tension
avoiding} protocol (\secref{TensionAvoidance}). We describe how tension
avoiding protocols carefully enlarge quorum sizes to sidestep the tension. We
also explain how Basic EPaxos~\cite{moraru2013there} and
Atlas~\cite{enes2020state} can be expressed as optimized variants of Unanimous
\BPaxos{}.

Fourth, we introduce \textbf{Majority Commit \BPaxos{}}, a \emph{tension
resolving} protocol (\secref{TensionResolution}). We describe how tension
resolving protocols perform detective work to resolve the tension without
enlarging quorum sizes.  We also discuss the relationship between Majority
Commit \BPaxos{} and other tension resolving protocols like
EPaxos~\cite{moraru2013proof} and Caesar~\cite{arun2017speeding}.

In summary, we make the following contributions.
\begin{itemize}
  \item
    We explain generalized multi-leader protocols carefully and thoroughly,
    bringing clarity to an otherwise dense area of popular research.

  \item
    We present four new generalized multi-leader state machine replication
    protocols: Simple \BPaxos{}, Fast \BPaxos{}, Unanimous \BPaxos{}, and
    Majority Commit \BPaxos{}.

  \item
    We identify a fundamental tension between agreeing on commands and ordering
    commands and use this insight to taxonomize generalized multi-leader
    protocols into those that avoid the tension and those that resolve it.
\end{itemize}
