\section{Majority Commit BPaxos}\seclabel{MajorityCommitBPaxos}
\subsection{The Protocol}
Majority Commit BPaxos is identical to Deadlock BPaxos except for the following
modifications.
%
First, we change the condition under which a value is considered chosen on the
fast path.  Deadlock BPaxos considered a value $v = (x, \deps{I})$ chosen on
the fast path if a fast quorum $\FastQuorum$ of acceptors voted for $v$ in
round $0$.  Majority Commit BPaxos, on the other hand, considers a value $v =
(x, \deps{I})$ chosen on the fast path if a fast quorum $\FastQuorum$ of
acceptors voted for $v$ in round $0$ \emph{and} for every instance $I' \in
\deps{I}$, there exists a quorum $\Quorum' \subseteq \FastQuorum$ of $f + 1$
acceptors that knew $I'$ was chosen at the time of voting for $v$.
%
Second, when a Fast Paxos acceptor $a_i$ sends a phase 2b vote in round $0$ for
value $v = (x, \deps{I})$, $a_i$ also includes the subset of instances in
$\deps{I}$ that $a_i$ knows are chosen, as well as the values chosen in these
instances.
%
Third, Majority Commit BPaxos nodes execute \algoref{DeadlockBPaxos} but with
the lines of code shown in \algoref{MajorityCommitBPaxos} inserted after line
4.

\begin{algorithm}[ht]
  \caption{Majority Commit BPaxos recovery of instance $I$ by $b_i$}%
  \algolabel{MajorityCommitBPaxos}
  \begin{algorithmic}[1]
    \makeatletter
    \setcounter{ALG@line}{4}
    \makeatother
    \State $\mathcal{A} \gets$ the set of acceptors in $\Quorum$ that voted for
           $v$ in round $0$
    \If{$\exists I' \in \deps{I}$ such that no acceptor in $\mathcal{A}$
        knows that $I'$ is chosen}
      \State send any value satisfying \invref{PrunedDependencies}
    \EndIf{}
    \If{$b_i$ was recovering $I'$ and deferred to the recovery of $I$}
      \If{$I' \in \deps{I}$}
        \State abort recovery of $I'$; $I'$ has already been chosen
      \Else{}
        \State in instance $I'$, send any value satisfying
               \invref{PrunedDependencies}
      \EndIf{}
    \EndIf{}
  \end{algorithmic}
\end{algorithm}

We now explain \algoref{MajorityCommitBPaxos}. On line 5, $b_i$ computes the
subset $\mathcal{A} \subseteq \Quorum$ of acceptors that voted for $v$ in round
$0$. On line 6, $b_i$ determines whether there exists some instance $I' \in
\deps{I}$ such that no acceptor in $\mathcal{A}$ knows that $I'$ is chosen. If
such an $I'$ exists, then $v$ was not chosen in round $0$. To see why, assume
for contradiction that $v$ was chosen in round $0$. Then, there exists some
fast quorum $\FastQuorum$ of acceptors that voted for $v$ in round $0$, and
there exists some quorum $\Quorum' \subseteq \FastQuorum$ of acceptors that
know $I'$ has been chosen. However, any two quorums intersect, but no acceptor
in $\Quorum$ both voted for $v$ in round $0$ and knows that $I'$ was chosen.
This is a contradiction. Thus, $b_i$ is free to propose any value satisfying
\invref{PrunedDependencies} on line 7.

Next, it's possible that $b_i$ was previously recovering instance $I'$ with
value $v' = (x', \deps{I'})$ and executed line 11 of \algoref{DeadlockBPaxos},
deferring the recovery of instance $I'$ until after the recovery of instance
$I$.
%
If so, $b_i$ continues on line 9. If $I' \in \deps{I}$, then some acceptor $a_j
\in \mathcal{A}$ knows that $I'$ is chosen. Thus, $b_i$ can abort the recovery
of instance $I'$ and retrieve the chosen value directly from $a_j$.
%
Otherwise, $I' \notin \deps{I}$. In this case, no value was chosen in round $0$
of instance $I'$, so $b_i$ is free to propose any value satisfying
\invref{PrunedDependencies} in instance $I'$. Here's why. $I' \notin \deps{I}$,
so every dependency service node co-located with an acceptor in $\mathcal{A}$
processed $I$ before $I'$. $|\mathcal{A}| \geq \QuorumMajoritySize$, so there
strictly fewer than $f + \QuorumMajoritySize$ remaining dependency service
nodes that could have processed $I'$ before $I$.  If $b_i$ was recovering
instance $I'$ but deferred to the recovery of instance $I$ on line 11 of
\algoref{DeadlockBPaxos}, then $I \notin \deps{I'}$.  In order for $I'$ to have
been chosen in round $0$ with $I \notin \deps{I'}$, it requires that at least
$f + \QuorumMajoritySize$ dependency service nodes processed $I'$ before $I$,
which we just concluded is impossible. Thus, $I'$ was not chosen in round $0$.

\subsection{Safety and Deadlock Freeness}
Majority Commit BPaxos' proof of safety is a natural extension of Deadlock
BPaxos's proof of safety. Majority Commit BPaxos maintains
\invref{ConsensusInvariant} by implementing Fast Paxos, and it maintains
\invref{ConflictInvariant} by maintaining \invref{PrunedDependencies}.
%
Moreover, Majority Commit BPaxos is deadlock free. If a Majority Commit BPaxos
node $b_i$ is recovering instance $I'$ and defers to the recovery of instance
$I$, then either $b_i$ will recover $I$ using line 7 of
\algoref{MajorityCommitBPaxos} or $b_i$ will recover $I'$ using line 10 or line
12 of \algoref{MajorityCommitBPaxos}. In either case, any potential deadlock is
avoided.
