\section{Unsafe BPaxos}\appendixlabel{UnsafeBPaxosAppendix}
We now explain why Unsafe BPaxos is unsafe. Consider an Unsafe BPaxos
deployment with $f = 2$ and $n = 2f + 1 = 5$.
%
Assume Unsafe BPaxos node $b_1$ receives command $x$ from a client and
Unsafe BPaxos $b_2$ receives a conflicting command $y$ from a client. $b_1$
sends $(I_x, x)$ to the dependency service, and $b_2$ sends $(I_y, y)$ to the
dependency service. Dependency service nodes $d_1$ and $d_2$ receive $(I_x, x)$
and propose $(x, \emptyset)$ in instance $I_x$ to Fast Paxos acceptors $a_1$
and $a_2$. Similar, $d_4$ and $d_5$ receive $(I_y, y)$ and propose $(y,
\emptyset)$ in instance $I_y$ to acceptors $a_4$ and $a_5$. Then, $b_1$ and
$b_2$ crash and all other messages are dropped. Unsafe BPaxos node $b_3$
then attempts to recover $I_x$. In phase 1 of Fast Paxos, $b_3$ receives phase
1b messages from $a_1$, $a_2$, and $a_3$. Because $a_1$ and $a_2$ both voted
for the value $(x, \emptyset{})$ in round $0$, $b_3$ is forced to propose the
value $(x, \emptyset)$ in phase 2 (\lineref{FastPaxosCase2} of
\algoref{FastPaxos}). Assume this value gets chosen. Then, $b_3$ recovers
$I_y$. In phase 1 of Fast Paxos, $b_3$ receives phase 1b messages from $a_3$,
$a_4$, and $a_5$. $a_4$ and $a_5$ both voted for the value $(y, \emptyset{})$
in round $0$, so $b_3$ is forced to propose the value $(y, \emptyset)$ in phase
2 (\lineref{FastPaxosCase2} of \algoref{FastPaxos}). Assume this value gets
chosen. Now, instances $I_x$ and $I_y$ have both been chosen with conflicting
commands $x$ and $y$, but neither instance depends on the other. This violates
\invref{ConflictInvariant}, and can result in two replicas executing
conflicting commands in different orders.

Unsafe BPaxos illustrates a fundamental tension between preserving
\invref{ConsensusInvariant} and preserving \invref{ConflictInvariant}.
Maintaining \invref{ConsensusInvariant} in isolation is easy (e.g., use Paxos),
and maintaining \invref{ConflictInvariant} in isolation is also easy (e.g., use
the dependency service). But, maintaining both invariants simultaneously is
tricky. There are situations, like the one in our example above, where a BPaxos
node is forced to propose a particular value $(x, \deps{I_x})$ in order to
preserve \invref{ConsensusInvariant} (e.g., because the value may have been
chosen in an earlier Fast Paxos round) and simultaneously forced \emph{not} to
propose the value in order to preserve \invref{ConflictInvariant} (e.g.,
because $\deps{I_x}$ may not have been computed by the dependency service).
%
In the next couple of sections, we introduce a handful of BPaxos protocols,
each of which runs into this fundamental tension. The distinguishing feature of
every protocol is how it resolves the tension.
