\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{jsys}

% Disable a warning related to \nonfrenchspacing
\microtypecontext{spacing=nonfrench}

\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{bpaxos}
\usepackage{enumitem}
\usepackage{pervasives}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}

% Reviewer quotes.
\definecolor{ReviewerDarkGray}{HTML}{37474F}
\newenvironment{reviewerquote}
{\list{}{\leftmargin=\parindent\rightmargin=0in}\item[] \itshape \color{ReviewerDarkGray}}%
{\endlist}

\newtoggle{showfeedback}
% \toggletrue{showfeedback}
\togglefalse{showfeedback}
\NewEnviron{feedback}{
  \iftoggle{showfeedback}{
    \begin{itemize}[leftmargin=*]
      \item[] \itshape \textcolor{ReviewerDarkGray}{\BODY}
    \end{itemize}
  }{}
}

\begin{document}
\title{Response to Reviewer Comments}

\maketitle

%% Disable header and footer on the from page.
\thispagestyle{empty}

We'd first like to thank the reviewers for all of their feedback. Below, we've
included our response to the reviewer feedback and our revised paper. Changes
to the paper are shown in \markrevisions{pink}.
% The major changes include...

\section*{Reviewer 1}
\begin{reviewerquote}
  Weakness

  Although the paper targets to be a pedagogical paper, it is not easy to read.
  Cleaning up writing may be necessary.  Paper often gets not precise. Clear
  assumptions and thorough correctness arguments are missing.  I think this
  paper has the potential to be a good tutorial to dependency graph-based
  consensus protocols. (I might change the title. I think this paper is
  specifically on dependency graph-based Paxos protocols, not for general
  multi-leader consensus protocols.) However, it wasn't a paper that reads
  smoothly, especially since this paper presents many different protocols. As
  it is targetting to be a pedagogical paper, I recommend authors to spend time
  on that. In the detailed comments below, I added a few suggestions.

  Also, the paper gets loose on preciseness from time to time. I understand
  that it might be impossible to present proofs to all protocols presented. But
  it sometimes felt too loose to wrap my mind about correctness.

  Finally, I felt the paper wanders among various PPaxos \& other prior work. I
  wonder if setting up a goal upfront would be helpful. It might be useful to
  set the goal of understanding EPaxos like protocols and why people want them.
\end{reviewerquote}

\begin{reviewerquote}
  Page 1.

  \begin{itemize}
    \item
      ``generalized [13, 18]'' $\Rightarrow$ Is it a widely accepted term?
      Personally, I didn't really see that term outside of generalized Paxos.
  \end{itemize}
\end{reviewerquote}

Atlas~\cite{enes2020state}, Caesar~\cite{arun2017speeding}, and a handful of
other papers in the space~\cite{sutra2011fast, losa2016brief} use the term
``generalized'', so we decided to use it as well to be consistent with these
papers.

\begin{reviewerquote}
  \begin{itemize}
    \item
      In the introduction, I would clarify that this paper's scope is on
      dependency graph-based Paxos protocols. It
  \end{itemize}
\end{reviewerquote}

We added a sentence to the introduction clarifying that the paper focuses on
protocols based on dependency graphs.

\begin{reviewerquote}
  Page 2.

  \begin{itemize}
    \item
      ``For example, with two proposers... the leader of the round'' seems not
      important.
  \end{itemize}
\end{reviewerquote}

We removed the example from the paper. We left the sentence clarifying our
usage of the word ``leader'' though, as we use this a lot throughout the paper.

\begin{reviewerquote}
  \begin{itemize}
    \item
      ``chosen'' at the end is not defined. It seems different from ``chosen''
      in page 4. I would clarify ``chosen'' and ``voted'' in page 3.
  \end{itemize}
\end{reviewerquote}

We edited Section 2.2 to make the definition of chosen more clear. Formally, a
value is considered chosen when a majority of the acceptors vote for the value
in the same round.

\begin{reviewerquote}
  Page 3.

  \begin{itemize}
    \item
      Figure 1 added little value to me. Except ``number of leaders'', terms in
      other boxes are not yet defined. Also, it covers many protocols not
      discussed in the paper.
  \end{itemize}
\end{reviewerquote}

We moved Figure 1 to Section 10 at the end of the paper; it's now Figure 15.
Now, all of the terms in the figure are defined when the figure is introduced.

\begin{reviewerquote}
  \begin{itemize}
    \item
      Ditto for Table 1. ``$\maj{}$'' is not defined either.
  \end{itemize}
\end{reviewerquote}

We also moved Table 1 to Section 10 at the end of the paper; it's now Table 2.
As with Figure 1, all of the terms in the table are now defined when the table
is introduced.

\begin{reviewerquote}
  \begin{itemize}
    \item
      I add details to Figure 3. The current version only shows how many
      message delays are required. I think a figure with terms $(vr, vv,
      \PhaseIA{i}, \text{etc})$ used in the text will be very helpful.
  \end{itemize}
\end{reviewerquote}

We added the terms you requested to the figure.

\begin{reviewerquote}
  \begin{itemize}
    \item
      Meaning of ``$v$'' is mixed. Sometimes ``$v$'' means the value given by
      the client. Sometimes it's not.
  \end{itemize}

  Page 4.

  \begin{itemize}
    \item
      If any message gets ignored, what does a proposer do?
  \end{itemize}

  \begin{itemize}
    \item
      Can you elaborate on how replicas work? Why is it safe, and how to ensure
      consistency?
  \end{itemize}

  Page 5.

  \begin{itemize}
    \item
      Need an explanation on directed edges. There is no introduction of the
      meaning of direction. The definition of dependency explained so far
      doesn't pose any order.
  \end{itemize}

  \begin{itemize}
    \item
      3.2, ``Replicas execute conflict graphs in reverse topological order''
      $\Rightarrow$ why? what goes wrong if executed in topological order? I
      could imagine it's like EPaxos, but unclear. I wish to see some
      explanation based on the ultimate goal.
  \end{itemize}

  \begin{itemize}
    \item
      3.3 ``a log one entry'' $\Rightarrow$ typo?
  \end{itemize}

  Page 7.

  \begin{itemize}
    \item
      ``$v_x = (p_i, m)$'' $\Rightarrow$ $v_x$, whose vertex id is $(p_i, m)$.
      Mixing command, vertex, and vertex id was a bit confusing. For example,
      ``globally unique vertex id $v_x = (p_i, m)$'' but later in text, $v_x$
      means vertex, not id.
  \end{itemize}

  Page 8.

  \begin{itemize}
    \item
      Texts for figure 11 are difficult to read. Maybe move them to caption.
  \end{itemize}

  \begin{itemize}
    \item
      Here, you introduce the timer. I think Paxos protocols generally don't
      require a synchronous network. I think this new requirement should be
      declared earlier.
  \end{itemize}

  \begin{itemize}
    \item
      Separate client request (command) from vertex id. If not, it seems that
      clients won't be able to retry after the vertex gets noop.
  \end{itemize}

  Page 9.

  \begin{itemize}
    \item
      4.5 doesn't prove that the two invariants guarantee linearizability or
      something similar. I would like to see the correctness argument for SMR,
      not just to the two invariants.
  \end{itemize}

  \begin{itemize}
    \item
      ``commit time'' is typically defined as the time period between a client
      sends a request and the client knows it's committed. Execution is not a
      requirement. It's an important distinction for EPaxos, whose commit
      latency is different from execution latency (which may be longer than
      commit latency).
  \end{itemize}

  \begin{itemize}
    \item
      I don't understand how EPaxos leverages FastPaxos. Also, EPaxos's key
      advantage is optimistic two wide-area message delays. More explanation is
      needed here on the delicate need of FastPaxos \& why 4 message delays are
      considered ``improvement'' despite MultiPaxos also only needs 4 message
      delays. I guess adding multi-leader requires more message delays, which
      could be removed by ``FastPaxos''.
  \end{itemize}

  Page 10.

  \begin{itemize}
    \item
      Why use both learner and replica? Any difference?
  \end{itemize}

  \begin{itemize}
    \item
      I feel ``$\maj{}$'' is awkward. Why not just use an equation? Also, I
      don't understand why $\maj{}$ needs to be defined separately for odd $n$
      and even $n$.
  \end{itemize}

  \begin{itemize}
    \item
      If you used the commit time defined earlier, it can get more than 4. It
      must wait for dependent commands.
  \end{itemize}

  \begin{itemize}
    \item
      6.2 Recovery $\Rightarrow$ the term recovery is overloaded. Here, it's
      just adding noop. In other places, it commits with real command.
  \end{itemize}

  \begin{itemize}
    \item
      In the bullet starting with ``In Figure 14b,'' $\Rightarrow$ remind
      readers about Paxos protocol. I think that's necessary to understand why
      ``so is forced to propose $(x, p_i)$''.
  \end{itemize}

  \begin{itemize}
    \item
      I initially didn't understand why proposer $p_2$ doesn't send the request
      to $d_1$, $d_2$, $\ldots$. It felt more natural to do that. But I guess
      you strictly abided by the Paxos protocol.
  \end{itemize}

  Page 14.

  \begin{itemize}
    \item
      7.1 $\Rightarrow$ first paragraph is complicated. Renaming phases and
      describing what they do would be helpful.
  \end{itemize}

  Page 15.

  \begin{itemize}
    \item
      Does Unanimous PPaxos modify the ``Paxos'' protocol part? Can you explain
      why replicas will still converge and agree?
  \end{itemize}
\end{reviewerquote}

\section*{Reviewer 2}
\begin{reviewerquote}
  Authors set a goal to encourage an industry adoption of the multi-leader
  protocols so I look and evaluate the paper from a position of a practitioner
  trying to build a multi-leader consensus based system. The explanation of
  fast paxos and multi-leader protocols is clear and it helped me to develop an
  intuition for the protocols sufficient for building a prototype.

  Overall it is a good paper: it not only refines the explanations but also
  introduces an abstraction helping to reason about the multi-leader consensus
  protocols (Simple PPaxos) and describes a new optimization / multi leader
  protocol (Majority Commit PPaxos). But the paper lacks several important
  characteristics necessary for building a production system:

  \begin{itemize}
    \item
      a reconfiguration routine
    \item
      information on the retrieving data from the state machines
    \item
      measurable benefits of switching to from Raft/MultiPaxos to the
      multi-leader protocols
  \end{itemize}

  Measurable benefits of switching to multi-leader protocols

  Multi-leader consensus protocols are complex and it's harder to implement
  them than their leadered counterparts but in theory they have better
  performance. As a practitioner I wonder if the benefits are worth the
  complexity but the paper doesn't give me an answer.

  The authors advocate that the major benefit comes as increased throughput (in
  Raft/Paxos a leader sends and receives disproportionately more messages). I
  recommend to include either experiment data or an estimation of the upper
  limits of throughput for both systems.

  E.g. napkin math shows that the throughput of multi-leader consensus is
  $(2f+1)^2/(4f+1)$ times better than leadered where $f$ is max allowed numbers
  of failures, proposer, acceptors and learners are colocated, it uses $f+1$
  sized quorums and the command's payload is way more bigger than the result of
  command execution.

  Retrieving data from the state machines

  Usually production Paxos-based systems depend on bypassing the replication
  protocol for serving reads but the paper doesn't contain any information on
  how to do it. For example the leader-based protocols have well understood
  models for lightweight reads:

  to read from a leader after waiting for a round of the heartbeats or after
  checking that its lease doesn't expire to read from a combination of the
  followers using methods from ``Linearizable Quorum Reads in Paxos'' or
  ``Paxos quorum leases: Fast reads without sacrificing writes'' papers to
  accept a possibility of staleness and to read from any node But those models
  aren't applicable to the multi-leader case: it lacks the leader/follower
  roles and local reads lead not only to staleness but to observing
  incompatible histories too caused by the reordering of the commuting
  commands.

  Reconfiguration

  The last important missing part is reconfiguration. It's crucial from a
  practical perspective. Without reconfiguration there is no way to replace the
  failed nodes and since failures are inevitable then eventually the
  multi-leader consensus based systems lose a majority and become unavailable.

  One of the reasons Raft got so widespread in the industry is an included
  reconfiguration protocol (joined consensus) so with the declared goal I
  recommend to update the paper to include the reconfiguration sub-protocol.

  Other (minor) issues

  ``This paper assumes that at most $f$ machines will fail'' - from a practical
  standpoint it's impossible to guarantee; it's better to mention what happens
  then this threshold is passed if consistency or availability becomes
  violated.

  ``To reach is consensus on a value, an execution of Paxos is divided into a
  number of integer value rounds'' - it's a clever idea to pre-assign a set of
  ballot numbers to the machines but using a term ``round'' both for a ballot
  number and an act of communication is confusing. It may create a false
  impression that it's impossible to skip ballot numbers. Maybe it'd be better
  to mention that rounds also known as ballot numbers (paxos) or term (raft)
  and add a reference to ``Paxos vs Raft: Have we reached consensus on
  distributed consensus?'' to use it as a map between different protocols.
\end{reviewerquote}

\section*{Reviewer 3}
\begin{reviewerquote}
  Summary:

  This paper aims to explain the space of multi-leader consensus protocols. The
  key advantage of adopting multi-leader protocols is that they facilitate
  out-of-order execution of commutating. In specific, if two commands do not
  conflict, then these protocols allow replicas to execute them in their
  desired order. To explain this space, the paper starts with a primer on SMR
  and Paxos, following which it explains MultiPaxos and defines the notion of
  conflict graphs. These conflict graphs are essential to the design of these
  multi-leader protocols as they specify the commands that can commute. To
  explain how existing multi-leader protocols work, authors lay down two key
  invariants: consensus invariant and dependency invariant. Next, authors
  gradually lay down the design of existing multi-leader protocols by
  presenting four of their variants: Simple PPaxos, Fast PPaxos, Unanimous
  PPaxos, and Majority Commit PPaxos. The paper also shows how these variants
  can be mapped to existing protocols like Basic Epaxos, Atlas, and so on.

  Pros:

  \begin{enumerate}
    \setcounter{enumi}{0}
    \item
      I think this paper is a timely addition to this space and has potential
      to become a paper like Raft that cleanly explains the domain of
      multi-leader protocols. With so many multi-leader protocols, a survey
      paper like this can be useful for both pedagogical and engineering
      purposes.
    \item
      The first seven pages of this paper are really fun to read and
      insightful. The authors have laid down the framework very well and kudos
      to them for the nice description. To motivate the spectrum, authors have
      started by explaining even Paxos, and as a reader, this really helps set
      up the pace.
    \item
      I found the paper as a whole to be very descriptive. For each protocol,
      authors have presented the algorithm, an accompanying figure, and an
      example. This helps inn visualizing the problem at hand. Further, the
      story is finely knit. Authors explain why Simple PPaxos is inefficient
      and then show how it can be improved by Fast PPaxos. However, Fast PPaxos
      is unsafe, so there are two approaches to follow: Unanimous PPaxos or
      Majority Commit PPaxos.
  \end{enumerate}

  Cons:

  \begin{enumerate}
    \setcounter{enumi}{0}
    \item
      I strongly believe that this paper is missing some evaluation. It will be
      really helpful if the authors can illustrate the performance of the four
      protocols they have proposed against some of the protocols like Epaox,
      Atlas and so on. Although the aim of this paper is not to evaluate
      performance of multi-leader protocols, graphs that illustrate the
      performance of the proposed variants can help understand various design
      choices, such as which scheme is better Tension Avoidance or Tension
      Resolution.
    \item
      This paper needs some more work to clarify the different proposed
      protocols. Further, for several sections, there is a sudden switch from
      normal phase to recovery phase, which makes it difficult to understand. I
      explain most of my concerns next:
  \end{enumerate}

  Major comments:

  \begin{enumerate}
    \setcounter{enumi}{0}
    \item
      While explaining the recovery for Simple PPaxos, the paper mentions
      "unchosen vertex" at the start of Section 4.4. It will be helpful if the
      paper explains what is meant by unchosen here. Further, it is important
      to re-iterate how can a vertex be unchosen.
    \item
      I am unsure why at line 21 of Algorithm 1, the propose is sending to at
      least $f+1$ acceptors instead of all acceptors. As $f$ of the acceptors
      may crash, so sending to all acceptors in important.
    \item
      It is unclear to me what is the benefit of waiting for messages from all
      the $2f+1$ acceptors in line 1, Algorithm 3. The paper should explain what
      will be the issue if the proposer only waits for messages from $f +
      \maj{f+1}$ acceptors. I understand that the change at line helps to resolve
      the safety bug of Fast PPaxos but not sure why change at line 1
      important. Moreover, asking Unanimous Paxos's proposer to wait for
      messages from all the $2f+1$ acceptors can hurt its liveness even if a
      single proposer fails. In such a case, how will the proposer recover the
      state. The paper needs to clarify this scenario.
    \item
      I think there is a typo for lines 6-7 in Algorithm 3. They are not
      identical to the lines is Algorithm 1.
    \item
      While explaining the Basic Epaxos optimization on Step 7, page 16, the
      paper states that ``pi receives $2f$ votes for... including a vote from
      $d_i$''.  I am not sure when did $d_i$ sent a vote to $p_i$. In the
      presented algorithm, $d_i$ only communicates with other $d_j$. This step
      needs to be clarified.
    \item
      It is unclear what is happening from steps (7) and (8) on Basic Epaxos.
      So, when $p_i$ receives $2f$ votes it sends $v$ to $a_i$. Then, $a_i$
      records $v$ as chosen and sends back to $p_i$. Then $p_i$ sends to all
      other ancestors, and waits for acknowledgments. Why can't these steps be
      combined? Further, to reduce from $2f+1$ to $2f$, the Basic Epaxos
      optimization requires more network communication--from proposer to all
      acceptors and then acknowledgments from acceptors to the proposer. This
      needs to be clarified.
  \end{enumerate}

  Minor comments:

  \begin{enumerate}
  \item
    On Page 6, the paper states Consensus Invariants, which uses the terms $(x,
    \deps{x})$. These terms have not been mentioned anytime before in this
    subsection. Although it can be understood by further reading, it will be
    helpful to the reader if explained before.
  \item
    It will be helpful if the paper either uses the term replicas or learners.
    The paper switches between the terms, which makes it often confusing.
  \item
    The following line ``We consider what happens when not every...'' on page
    11, left hand side column are confusing the reader, as I expected that
    following this line the solution will be presented, but that happens much
    later in this section. Either consider removing this line or re-write it a
    little.
  \item
    It is worth separating the recovery algorithm and not receiving enough
    matching votes (in Fast Paxos section) with suitable titles to improve
    readability.
  \end{enumerate}
\end{reviewerquote}

\section*{Reviewer 4}
\begin{reviewerquote}
  I enjoyed reading this tutorial paper on generalized multi-leader state
  machine replication. The paper gives a very simple protocol that nicely
  captures the essence of such protocols. The paper then presents an unsafe
  protocol to set the stage for why advanced protocols work the way they do in
  the literature. I am usually skeptical about deriving insights from unsafe
  protocols, but this paper does a good job with this approach.

  However, I am not entirely certain what went wrong for Fast PPaxos. The paper
  claims that it is due to a dilemma between reaching consensus and preserving
  dependency. But it seems to me that the bug is a result of multi-leader
  rather than dependency. It seems that the bug could be avoided by simply
  abandoning concurrent leaders. The first approach to fix this bug, increasing
  quorum size to 2f+1, also does not seem related to dependency; rather, it
  seems to be avoiding a tie between concurrent leaders. I would like to see
  more discussion on the root cause of the problem, and how exactly it is
  fixed. This is very important for the paper since at some point the authors
  claim that most of the design complexity comes from generalized as opposed to
  multi-leader.

  Other than that, I have a few minor suggestions in terms of presentation.

  \begin{itemize}
    \item
      When describing Fast Paxos, the authors use proposals, acceptors, and
      learners; but in Simple or Fast PPaxos, the authors use proposals,
      acceptors, and replicas. I suggest the authors stick to one convention,
      preferrably learners.
  \end{itemize}

  \begin{itemize}
    \item
      Page 11 right column touched on why the threshold must be > 1.5f. While
      this may be obvious for experts, it is helpful to elaborate on this in a
      tutorial paper for non-expert readers.
  \end{itemize}

  \begin{itemize}
    \item
      Section 2 says that messages can be dropped, and indeed Section 6.3
      considers a situation that messages to d3, d4, d5 are dropped. However,
      the default model of Paxos does not allow message drop. Otherwise, one
      runs into the two-general impossibility. In this particular example in
      Section 6.3, can we simply have p1 crash before it sends messages to d3,
      d4,and d5? If this is sufficient to break Fast PPaxos, then I suggest
      removing message drop from the model. If this is not sufficient, the
      authors need to carefully define the model regarding message drop, and
      explain why the two general's impossibility does not apply.
  \end{itemize}

  Overall, I like this tutorial paper and would support its acceptance as a
  tutorial paper. However, it is currently submitted as an SoK paper. If judged
  as an SoK paper, I am less supportive. The paper mentions that there are few
  generalized multi-leader systems, and indeed it only reviews 3 papers in
  detail (EPaxos, Caesar, Atlas). The related work section mentions a few other
  works very briefly and they are not closely related. As such, I feel the area
  of generalized multi-leader state machine replication is still relatively
  young and small, and has not reached a stage of needing an SoK paper. This is
  not a criticism of the paper. I think the authors are well aware that their
  paper is more of a tutorial than SoK. The onus is on the editorial board to
  accept this paper under an appropriate category.
\end{reviewerquote}

\bibliographystyle{plain}
\bibliography{references}

% \clearpage

\end{document}
