\section{Bipartisan Paxos}\seclabel{BipartisanPaxos}
In this section, we describe \defword{Bipartisan Paxos}, or \defword{BPaxos}.
BPaxos is designed to be as simple to understand as possible, even at the cost
of performance. Other variants of BPaxos that we'll see later (i.e.\ Unanimous
BPaxos and Fast BPaxos) improve on BPaxos' performance.

\paragraph{Overview}
MultiPaxos allows a set of nodes to agree on a totally ordered sequence of
state machine commands. This is illustrated in \figref{MultiPaxosCartoon}.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \node[square] (a) at (0, 0) {\texttt{x=1}};
    \node[square, right=-1pt of a] (b) {\texttt{y=1}};
    \node[square, right=-1pt of b] (c) {\texttt{y+=x}};
    \node[square, right=-1pt of c] (d) {\texttt{y+=1}};
    \node[square, right=-1pt of d] (e) {\texttt{x=2\\y=2}};
    \node[square, right=-1pt of e] (f) {};
    \node[square, right=-1pt of f] (g) {};
    \node[right=0in of g] (h) {\ldots};

    \foreach \label/\i in {a/0, b/1, c/2, d/3, e/4, f/5, g/6} {%
      \node[below=0in of \label] {\i};
    }
  \end{tikzpicture}
  \caption{MultiPaxos}\figlabel{MultiPaxosCartoon}
\end{figure}

While simple, agreeing on a totally ordered sequence of state machine commands
can be overly prescriptive. If two commands conflict (e.g., \texttt{x = 1} and
\texttt{x = 2}), then they \emph{do} need to be executed by every state machine
replica in the same order. But, if two commands do \emph{not} conflict (e.g.,
\texttt{x = 1} and \texttt{y = 1}), then they do \emph{not} need to be totally
ordered.  State machine replicas can execute them in either order.

EPaxos takes advantage of this fact and attempts to order commands only if they
conflict. To do so, it ditches the totally ordered sequence and instead agrees
on a directed graph of commands such that every pair of conflicting commands
have an edge between them. An example is illustrated in \figref{EPaxosCartoon}%
\footnote{%
  In reality, the graph would be the transitive closure of the one in
  \figref{EPaxosCartoon}, but we do not draw all edges to keep things simple
}.
EPaxos then executes this graph in reverse topological order one strongly
connected component at a time, executing commands within a component in an
arbitrary but deterministic order. For example, in \figref{EPaxosCartoon}, we
could execute commands in the order $R.1, R.2, S.1, Q.2, Q.1$.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}[scale=2.5]
    \node[square] (a) at (0, 1) {\texttt{x=1}};
    \node[square] (b) at (0, 0) {\texttt{y=1}};
    \node[square] (c) at (1, 1) {\texttt{y+=x}};
    \node[square] (d) at (1, 0) {\texttt{y+=1}};
    \node[square] (e) at (2, 0.5) {\texttt{x=2\\y=2}};
    \draw[ultra thick, -latex] (c) to (a);
    \draw[ultra thick, -latex] (c) to (b);
    \draw[ultra thick, -latex, bend right] (c) to (e);
    \draw[ultra thick, -latex] (d) to (b);
    \draw[ultra thick, -latex, bend right] (e) to (c);
    \draw[ultra thick, -latex] (e) to (d);

    \foreach \label/\i in {a/$R.1$, b/$R.2$, c/$Q.1$, d/$S.1$, e/$Q.2$} {%
      \node[above=0in of \label] {\i};
    }
  \end{tikzpicture}
  \caption{EPaxos}\figlabel{EPaxosCartoon}
\end{figure}

Every vertex in the graph has a unique name like $Q.1$ or $R.1$. EPaxos calls
these \defword{instances}. We can view a named vertex, the command inside the
vertex, and the vertex's outbound edges as a little gadget. For example,
\figref{EPaxosGadgets} shows gadgets for instances $R.2$, $Q.1$, and $Q.2$.
%
More formally, we can think of these gadgets as tuples like $(Q.1,
\texttt{y+=x}, \set{R.1, R.2, Q.2})$ where $Q.1$ is the instance, \texttt{y+=x}
is the command in the instance, and the set $\set{R.1, R.2, Q.2}$ is the set of
\defword{dependencies} of $Q.1$ (or of \texttt{y+=x} if $Q.1$ is clear from
context).

\begin{figure}[h]
  \centering

  \begin{subfigure}[b]{0.19\textwidth}
    \begin{tikzpicture}[scale=2.5]
      \node[square] (b) at (0, 0) {\texttt{y=1}};
      \foreach \label/\i in {b/$R.2$} {%
        \node[above=0in of \label] {\i};
      }
    \end{tikzpicture}
  \end{subfigure}
  %
  \begin{subfigure}[b]{0.49\textwidth}
    \begin{tikzpicture}[scale=2.5]
      \node[square] (a) at (0, 1) {};
      \node[square] (b) at (0, 0) {};
      \node[square] (c) at (1, 1) {\texttt{y+=x}};
      \node[square] (e) at (2, 0.5) {};
      \draw[ultra thick, -latex] (c) to (a);
      \draw[ultra thick, -latex] (c) to (b);
      \draw[ultra thick, -latex, bend right] (c) to (e);
      \foreach \label/\i in {a/$R.1$, b/$R.2$, c/$Q.1$, e/$Q.2$} {%
        \node[above=0in of \label] {\i};
      }
    \end{tikzpicture}
  \end{subfigure}
  %
  \begin{subfigure}[b]{0.29\textwidth}
    \begin{tikzpicture}[scale=2.5]
      \node[square] (c) at (1, 1) {};
      \node[square] (d) at (1, 0) {};
      \node[square] (e) at (2, 0.5) {\texttt{x=2\\y=2}};
      \draw[ultra thick, -latex, bend right] (e) to (c);
      \draw[ultra thick, -latex] (e) to (d);
      \foreach \label/\i in {c/$Q.1$, d/$S.1$, e/$Q.2$} {%
        \node[above=0in of \label] {\i};
      }
    \end{tikzpicture}
  \end{subfigure}

  \caption{EPaxos Gadgets}\figlabel{EPaxosGadgets}
\end{figure}

EPaxos nodes collectively construct the directed graph of commands by stitching
together a bunch of gadgets. More precisely, an EPaxos node $R$ proposes a
gadget for instance $R.i$ and attempts to get the gadget chosen. Once a gadget
is chosen, it is considered part of the directed graph and is eligible for
execution. EPaxos' correctness hinges on the following two key invariants
(later we'll see why these invariants ensure EPaxos' correctness).

\begin{boxedinvariant}\invlabel{GadgetsChosen}
  Once a gadget $(R.i, a, \deps{a})$ has been chosen, no other gadget can be
  chosen for instance $R.i$.
\end{boxedinvariant}

\begin{boxedinvariant}\invlabel{ConflictingGadgets}
  If two gadgets $(I_a, a, \deps{a})$ and $(I_b, b, \deps{b})$ are chosen and
  commands $a$ and $b$ conflict, then either $I_a \in \deps{b}$ or $I_b \in
  \deps{a}$ or both.
\end{boxedinvariant}

BPaxos, like EPaxos, also constructs a directed graph of commands and executes
them in reverse topological order one strongly connected component at a time.
In fact, BPaxos and EPaxos execute commands in 100\% the same way. BPaxos also
maintains the same two key invariants as EPaxos. BPaxos and EPaxos differ only
in how they construct the directed graph of commands.
%
BPaxos is illustrated in \figref{BPaxos}. A BPaxos instance has three main
components: an ordering service, a consensus service, and a set of BPaxos
nodes. The ordering service helps provide \invref{ConflictingGadgets}, the
consensus service helps provide \invref{GadgetsChosen}, and the BPaxos nodes
glue the two together. We explain each of these three components in turn.

{\input{figures/bpaxos.tex}}

\paragraph{Ordering Service}
The ordering service is responsible for computing the dependencies between
conflicting state machine commands. When a BPaxos node $R$ receives a state
machine command $a$ from a client, it chooses some instance $R.i$ for $a$.
Then, $R$ sends $a$ and $R.i$ to the ordering service. When the ordering
service receives a command $a$ for instance $R.i$, it replies with a tuple
$(R.i, a, \deps{a})$ where $\deps{a} = \set{I_1, \ldots, I_n}$ is $a$'s
dependencies.
%
The ordering service maintains the following invariant.

\begin{boxedinvariant}\invlabel{OrderingService}
If two conflicting commands $a$ and $b$ in instances $I_a$ and $I_b$ yield
responses $(I_a, a, \deps{a})$ and $(I_b, b, \deps{b})$ from the ordering
service, then either $I_a \in \deps{b}$ or $I_b \in \deps{a}$ (or both). That
is, if two conflicting commands are sent to the ordering service, at least one
is guaranteed to be a dependency of the other.
\end{boxedinvariant}

Note that the ordering service makes the following assumption. At most one
command can be sent in any given instance. BPaxos satisfies this invariant
because $R$ assigns every command $a$ to a unique instance $R.i$, and no other
EPaxos node contacts the ordering service for an instance $R.i$.

Implementing a fault tolerant ordering service is straightforward. We employ
$2f + 1$ ordering service nodes $o_{1}, \ldots, o_{2f + 1}$. When a BPaxos node
$R$ sends a command $a$ in instance $R.i$ to the ordering service, it sends the
command to all $2f + 1$ of the ordering service nodes. Every ordering service
node $o_i$ maintains a set $O_i = {(I_1, c_1), (I_2, c_2), \ldots}$ of all the
commands and corresponding instances that it has received. When node $o_i$
receives a command $a$ for instance $R.i$ from a BPaxos node, it atomically
performs two actions.
\begin{enumerate}
  \item
    $o_i$ replies to $R$ with the tuple $(R.i, a, \deps{a}_i)$ where
    $\deps{a}_i$ is the set of instances $I$ for which there exists a $(I, c)
    \in O_i$ such that $a$ and $c$ conflict. This is the set of instances that
    $o_i$ has previously received that contain a command that conflict with
    $a$.

  \item
    $o_i$ adds $(R.i, a)$ to $O_i$.
\end{enumerate}

When a BPaxos node receives replies $(R.i, a, \deps{a}_{i_1}), \ldots, (R.i, a,
\deps{a}_{i_{f+1}})$ from a quorum $Q_a$ of $f + 1$ ordering service nodes, it
takes $(R.i, a, \deps{a}_{i_1} \cup \ldots \cup \deps{a}_{i_{f+1}})$ to be the
response from the ordering service. That is, it computes $\deps{a}$ by taking
the union of $\deps{a}_i$ from a majority of the ordering service nodes.

To understand why this ordering service implementation maintains
\invref{OrderingService}, consider conflicting commands $a$ and $b$ in
instances $I_a$ and $I_b$. Assume $a$'s reply $(I_a, a, \deps{a})$ was formed
from a quorum $Q_a$ and $b$'s reply $(I_b, b, \deps{b})$ was formed from a
quorum $Q_b$. Any two quorums intersect, so $Q_a \cap Q_b$ is nonempty. Let
$o_i$ be an ordering service node in this intersection. $o_i$ either received
$a$ or $b$ first. If it received $a$ first, then $(I_a, a)$ is in $O_i$ when
$o_i$ processed command $b$, so $o_i$ includes $I_a$ in $\deps{b}_i$, so $I_a$
is in $\deps{b}$.  Symmetrically, if $o_i$ received $b$ first, then it includes
$I_b$ in $\deps{a}$.

\paragraph{Consensus Service}
We assume we have some set $p_1, \ldots, p_n$ of nodes that implement Plain
Jane consensus. A BPaxos node can propose to the consensus service that some
value $v$ be chosen in some instance $I$. The consensus service replies with
the value that has been chosen in instance $I$, which may or may not be $v$
depending on if there are concurrent proposers proposing to instance $i$. The
consensus service guarantees that for every instance $I$, at most one value is
ever chosen in $I$.

We can implement the consensus service with one Paxos instance for every BPaxos
instance, with consensus service nodes playing the role of Paxos acceptors and
the BPaxos nodes playing the role of Paxos proposers. Similar to MultiPaxos, we
can have each BPaxos node $R$ serve as the leader for every Paxos instance
$R.i$. Initially, $R$ runs phase 1 of Paxos for every instance $R.i$, so that
later, $R$ can get a command committed in instance $R.i$ in one round trip to
the consensus service (in the best case).

\paragraph{BPaxos Nodes}
We assume a fixed set $R_1, \ldots, R_{2f+1}$ of $2f + 1$ BPaxos nodes.
%
Clients sends state machine commands to BPaxos nodes to be executed by the
replicated state machine. When a BPaxos node $R$ receives a command $a$, it
sends the command to the ordering service in a previously unused instance $R.i$
and receives a reply $(R.i, a, \deps{a})$. $R$ then proposes the value $(a,
\deps{a})$ to the consensus service in instance $R.i$. The consensus service
then replies with some chosen value $(a', \deps{a'})$ (which is equal to $(a,
\deps{a})$ in the failure-free case). At this point, the gadget $(R.i, a',
\deps{a'})$ is considered chosen in the directed graph of state machine
commands and is eligible for execution. Node $R$ also informs the other BPaxos
nodes that the value $(a', \deps{a'})$ has been chosen.

As noted earlier, the execution of the commands in the directed graph is
identical to that of EPaxos. Committed commands are executed in reverse
topological order, one strongly connected component at a time. Within a
strongly connected component, BPaxos executes commands in an arbitrary but
deterministic order. Unlike with EPaxos, BPaxos does not annotate instances
with sequence numbers. Thus, BPaxos provides serializability instead of
linearizability. This is not fundamental. It just makes things a bit simpler.

It's possible that (1) a committed command $a$ depends on an instance $R.i$
that contains an uncommitted command, and (2) the BPaxos node $R$ that manages
the instance $R.i$ has crashed. If the instance $R.i$ remains forever
uncommitted, then the command $a$ will never be executed. To avoid this
liveness violation, if any BPaxos node $S$ notices that instance $R.i$ has been
uncommitted for some time, it can propose to the consensus service that the
value $(\noop, \emptyset)$ be chosen in instance $R.i$ where $\noop$ is a
command that doesn't affect the state machine and doesn't conflict with any
other command.

\paragraph{Correctness}
BPaxos maintains \invref{GadgetsChosen} trivially because it chooses commands
using consensus. The ordering service maintains \invref{OrderingService}, which
says that the dependencies computed for conflicting commands $a$ and $b$ will
have one in the other (or both). \invref{ConflictingGadgets} says that the
dependencies of conflicting \emph{chosen} commands $a$ and $b$ will have one in
the other (or both). Thus, BPaxos nodes maintain \invref{ConflictingGadgets} by
only proposing dependencies computed by the ordering service. BPaxos nodes also
propose $\noop$'s with empty dependencies, but since $\noop$s do not conflict
with any command, \invref{ConflictingGadgets} is satisfied trivially.

Now, we discuss why \invref{GadgetsChosen} and \invref{ConflictingGadgets}
ensure BPaxos' correctness. Fortunate for us, EPaxos maintains the same two
invariants, so we can leverage EPaxos' proof of correctness. Open up
\cite{moraru2013proof} and head to section 5.6, which contains proofs of
EPaxos' correctness.
\begin{itemize}
  \item
    \textbf{Theorem 1} says that EPaxos satisfies nontriviality. Clearly, so
    does BPaxos.

  \item
    \textbf{Lemma 1} is \invref{GadgetsChosen}. As we saw, BPaxos maintains
    this invariant.

  \item
    \textbf{Theorem 2} follows from Lemma 1.

  \item
    \textbf{Theorem 3} is trivial.

  \item
    \textbf{Theorem 4} states that if two conflicting commands are both
    committed, then they will be executed in the same order by every BPaxos
    node. This follows from \invref{ConflictingGadgets}. If two commands
    conflict, then one will depend on the other or both. If both commands end
    up in the same strongly connected component, they will be executed in the
    same deterministic order. And, if the commands end up in different strongly
    connected components, then one component is guaranteed to be ordered before
    the other, so the two commands are executed in reverse topological order.
\end{itemize}

